---
title: "Moovies Analysis"
output: html_document
---


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
```


# Analysis of movies- IMDB dataset

```{r,load_movies, warning=FALSE, message=FALSE}
movies <- read_csv(here::here("data", "movies.csv"))#import the data set of movies
glimpse(movies)#see the variables of movies
```

## Use your data import, inspection, and cleaning skills to answer the following:

- Produce a table with the count of movies by genre, ranked in descending order
```{r}
genres<-movies%>%group_by(genre)%>%count(sort=TRUE)
#group the data set by genre and count number of movies for each genre in descending order
genres
```

- Produce a table with the average gross earning and budget (`gross` and `budget`) by genre. Calculate a variable `return_on_budget` which shows how many $ did a movie make at the box office for each $ of its budget. Ranked genres by this `return_on_budget` in descending order
```{r}
means<-movies%>%group_by(genre)%>% summarise(avg_gross=mean(gross),avg_budget=mean(budget),return_on_budget=avg_gross/avg_budget,.groups = 'drop')%>%arrange(desc(return_on_budget))
#group the data set by genre, calculate the average gross earning and budget by genre and return_on_budget which equals to average gross earning/average budget, then use arrange() to range genres in descending order
means
```

- Produce a table that shows the top 15 directors who have created the highest gross revenue in the box office. Don't just show the total gross amount, but also the mean, median, and standard deviation per director.
```{r}
top15_director<- movies %>%group_by(director) %>%summarise(total_gross = sum(gross, na.rm=TRUE),avg_gross=mean(gross, na.rm=TRUE),med_gross=median(gross,na.rm=TRUE),sd_gross=sd(gross, na.rm=TRUE),count = n(),.groups = 'drop') %>%arrange(desc(total_gross),.groups = 'drop')%>%head(15)
#group movies by director, sum the total gross for each director, calculate the mean, median, sd of gross for each director, then use arrange() to range total gross in descending order and choose the top 15 values
top15_director
```

- Finally, ratings. Produce a table that describes how ratings are distributed by genre. We don't want just the mean, but also, min, max, median, SD and some kind of a histogram or density graph that visually shows how ratings are distributed. 
```{r}
ratings<-movies%>%group_by(genre)%>%summarise(avg_rating=mean(rating),min_rating=min(rating),max_rating=max(rating),med_rating=median(rating),sd_rating=sd(rating),.groups = 'drop')
#group movies by genre, calculate mean, min,max and SD of rating by using summarise()
ratings
ggplot(movies, aes(x = rating, colour = genre)) + geom_density() + labs(title = "Ratings distribution by genre", y = "Density of rating",  x = "Rating")
#draw the density plot of rating for each genre in one graph with different colors
ggplot(movies, aes(x = rating)) + geom_density() +facet_wrap(~genre)+labs(title = "Ratings distribution by genre", y = "Density of rating",  x = "Rating")
#draw the density plot of rating for each genre in different graphs for better visibility
```

## Use `ggplot` to answer the following


```{r, gross_on_fblikes}
ggplot(movies, aes(y=gross, x=cast_facebook_likes))+geom_point()+scale_x_log10()+scale_y_log10()
#draw the scatterplot between facebook likes and gross, scale this plot for better visibility
```
There seems to be some correlation between facebook likes and the gross box office, with higher facebook likes leading to more gross box office revenues. There are several outliers suggesting that facebook likes is not a great predictor of box office revenues. Y-axe is gross and X-axe is cast_facebook_likes.

- Examine the relationship between `gross` and `budget`. Produce a scatterplot and wri_logte one sentence discussing whether budget is likely to be a good predictor of how much money a movie will make at the box office.
```{r, gross_on_budget}
ggplot(movies, aes(y=gross, x=budget))+geom_point()+scale_x_log10()+scale_y_log10()
#draw the scatterplot between budget and gross, scale this plot for better visibility
```
Budget proves to be a better predictor of gross box office revenues with a higher correlation.It is likely that higher budgets result in better publicity such as higher facebook advertisements and better production levels such as special effects and wonderful scripts, therefore, budget is the good predictor.


```{r, gross_on_rating}
ggplot(movies, aes(y=gross, x=rating))+geom_point()+facet_wrap(~genre)+scale_y_log10()
#draw the scatterplot between rating and gross of each genre and scale this plot for better visibility
```
It is strange that movies with poor ratings can still make high revenues at the box office. There seems to be little correlation between the rating and money that a movie will make at the box office, therefore, IMDB ratings are not likely to be a good predictor of how much money a movie will make at the box office.
